---
title: "PREDICT_PARADIGM: Model Validity of PREDICT v2.2 and v2.3"
author: "Yuwei Wang"
date:  "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
  toc_depth: 3  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# READ ME
The R script is designed for our manuscript entitled the external validation and clinical utility assessment of PREDICT breast cancer prognostic model in young women with node-negative breast cancer. This R script assessed the performance of PREDICT v2.2 and PREDICT v2.3 in all women diagnosed with node-negative breast cancer. Using the script, we generated results for Table 3, Supplementary Table 1 and Supplementary Table 2.
# Load the libraries
```{r load the libraries, message=FALSE, warning=FALSE}
library(readxl)
library(readr)
library(dplyr)
library(plyr)
library(rms)
library(ggplot2)
library(survminer)
library(survival)
library(splines)
library(knitr)
library(table1)
library(boot)
library(tidyverse)
library(rsample)
library(timeROC)
library(pec)
library(riskRegression)
library(ggplot2)
```

# Load the imputed datasets
In the previous script, we performed multiple imputation and ran the PREDICT v2.2 and v2.3 algorithms in the imputed dataset. After runing the two models, we generated v2.2-predicted and v2.3-predicted all-cause mortality. The two datasets are named as dataimpute_v22 and dataimpute_v23, respecitively. 
```{r load two datasets, include=FALSE}
TNBC <- read_csv("/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/TNBC485_BRCA1statusupdate_in_August_2022.csv")
#table(TNBC$gBRCA1, TNBC$BRCA2, useNA = "always")
gBRCAmnr <- TNBC$palga_nr[which(TNBC$gBRCA1 == 1 | TNBC$BRCA2 == "gBRCA2m")]
gBRCAwtnr<- TNBC$palga_nr[which(TNBC$gBRCA1 == 0 & TNBC$BRCA2 %in% c("sBRCA2m", "BRCA2wt"))]

dataimpute_v22 <- read.csv(file='/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/dataimpute_v22_July.csv',stringsAsFactors=F,header=T,sep = ",") %>%
  dplyr::select(-X) %>%
  mutate(gBRCA12 = ifelse(palga_nr %in% gBRCAmnr, 1,
                          ifelse(palga_nr %in% gBRCAwtnr, 0, NA)))

dataimpute_v23 <- read.csv(file='/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/dataimpute_v23_July.csv',stringsAsFactors=F,header=T,sep = ",") %>%
  dplyr::select(-X) %>%
  mutate(gBRCA12 = ifelse(palga_nr %in% gBRCAmnr, 1,
                          ifelse(palga_nr %in% gBRCAwtnr, 0, NA)))
```

We excluded women under 25, as PREDICT doesn't provide predictions for women under 25.
```{r datasets for model v22}
## all women
df_v22 <- dataimpute_v22 %>%
  filter(T_LEEFT >= 25) %>%
  mutate(intrsub = ifelse((er == 1 | pr == 1) & her2 == 0, "HRposHER2neg",
       ifelse((er == 1| pr == 1) & her2 == 1, "HRposHER2pos",
              ifelse(er == 0 & pr == 0 & her2 == 0, "HRnegHER2neg",
                     ifelse(er == 0 & pr == 0 & her2 == 1, "HRnegHER2pos", NA)))))

## women with ER-positive tumors
df_v22_ERpos <- df_v22 %>%
        filter(ER.num > 0)
## women with ER-negative tumors
df_v22_ERneg <- df_v22 %>%
        filter(ER.num == 0)

## women of each subtype
df_v22_HRposHER2neg <- df_v22 %>%
  filter(intrsub == "HRposHER2neg")

df_v22_HRposHER2pos <- df_v22 %>%
  filter(intrsub == "HRposHER2pos")

df_v22_HRnegHER2pos <- df_v22 %>%
  filter(intrsub == "HRnegHER2pos")

df_v22_HRnegHER2neg <- df_v22 %>%
  filter(intrsub == "HRnegHER2neg")

### women of TNBC with gBRCA1/2m or gBRCA1/2wt
df_v22_HRnegHER2neg_gBRCA12m <- df_v22_HRnegHER2neg %>%
  filter(gBRCA12 == 1)
df_v22_HRnegHER2neg_gBRCA12wt <- df_v22_HRnegHER2neg %>%
  filter(gBRCA12 == 0)
```

```{r datasets for model v23}
## all women
df_v23 <- dataimpute_v23 %>%
  filter(T_LEEFT >= 25) %>%
  mutate(intrsub = ifelse((er == 1 | pr == 1) & her2 == 0, "HRposHER2neg",
       ifelse((er == 1| pr == 1) & her2 == 1, "HRposHER2pos",
              ifelse(er == 0 & pr == 0 & her2 == 0, "HRnegHER2neg",
                     ifelse(er == 0 & pr == 0 & her2 == 1, "HRnegHER2pos", NA)))))

## women with ER-positive tumors
df_v23_ERpos <- df_v23 %>%
        filter(ER.num > 0)
## women with ER-negative tumors
df_v23_ERneg <- df_v23 %>%
        filter(ER.num == 0)

## women of each subtype
df_v23_HRposHER2neg <- df_v23 %>%
  filter(intrsub == "HRposHER2neg")

df_v23_HRposHER2pos <- df_v23 %>%
  filter(intrsub == "HRposHER2pos")

df_v23_HRnegHER2pos <- df_v23 %>%
  filter(intrsub == "HRnegHER2pos")

df_v23_HRnegHER2neg <- df_v23 %>%
  filter(intrsub == "HRnegHER2neg")

### women of TNBC with gBRCA1/2m or gBRCA1/2wt
df_v23_HRnegHER2neg_gBRCA12m <- df_v23_HRnegHER2neg %>%
  filter(gBRCA12 == 1)
df_v23_HRnegHER2neg_gBRCA12wt <- df_v23_HRnegHER2neg %>%
  filter(gBRCA12 == 0)
```

# Load functions
We made several functions for pooling results of calibration and discrimination.

## Functions for pooling estimates 
pool_estimate can pool the results from estimation and the corresponding standard error of each imputed dataset, and was adapted from the pool_auc function from the package psfmi.This function follows Rubin's rule (note that the estimation is not log-transformed). The function returns the pooled estimation and 95% confidence interval.
```{r pooling function}
pool_estimate <- function(est, se, nimp) {
  RR_se <- function(est, se, nimp){ # a function to calculate total SE
    m <- nimp
    w_var <- mean(se^2) # within variance
    b_var <- var(est) # between variance
    t_var <- w_var + (1 + (1/m)) * b_var # total variance
    se_total <- sqrt(t_var) # total se
    r <- (1 + 1 / m) * (b_var / w_var)
    v <- (m - 1) * (1 + (1/r))^2 #degree of freedom
    t <- qt(0.975, v)
    res <- c(se_total, t)
    return(res)
  }
  
  est <- unlist(est)
  se <- unlist(se)
  if(length(est) != nimp)
    stop("Include c-statistic value for each imputed dataset")
  
  mean_est <- format(round(mean(est), 2), nsmall = 2)
  se_total <- RR_se(est, se, nimp=nimp)
  uCI <- format(round(mean(est) + (se_total[2]*se_total[1]), 2), nsmall = 2)
  lCI <- format(round(mean(est) - (se_total[2]*se_total[1]), 2), nsmall = 2)
  
  res <- paste0(mean_est, " (", lCI, "-", uCI, ")")

  return(res)
}
```

## Function for pooling data for calibration plot
pool_caliplot was only slightly different from pool_estimation. It returns the pooled estiamtion, lower 95% confidence interval and upper 95% confidence interval separately.
```{r pooling function for calibration plot}
pool_caliplot <- function(est, se, nimp) {
  RR_se <- function(est, se, nimp){ # a function to calculate total SE
    m <- nimp
    w_var <- mean(se^2) # within variance
    b_var <- var(est) # between variance
    t_var <- w_var + (1 + (1/m)) * b_var # total variance
    se_total <- sqrt(t_var) # total se
    r <- (1 + 1 / m) * (b_var / w_var)
    v <- (m - 1) * (1 + (1/r))^2 #degree of freedom
    t <- qt(0.975, v)
    res <- c(se_total, t)
    return(res)
  }
  
  est <- unlist(est)
  se <- unlist(se)
  if(length(est) != nimp)
    stop("Include c-statistic value for each imputed dataset")
  
  mean_est <- mean(est)
  se_total <- RR_se(est, se, nimp=nimp)
  uCI <- mean(est) + (se_total[2]*se_total[1])
  lCI <- mean(est) - (se_total[2]*se_total[1])
  
  res <- c(mean_est,  lCI,  uCI)

  return(res)
}
```

pool_caliplot_df returns a dataset that can be used to generate calibration plots. The function first divides women into quintiles based on their expected risks. Then the observed risks and its lower and upper confidence intervals are calculated, as well as the mean of the expected risks in each quintile.
```{r pooling data for plot}
pool_caliplot_df <- function(data, time, status, impvar, nimp, predmort, tmax) {
    res_obs <- matrix(NA, nimp, 5) 
    res_se <- matrix(NA, nimp, 5)
    res_exp <- matrix(NA, nimp, 5)

    for (i in 1:nimp) {

      df <- data[data[impvar] == i, ]
      
      q5 <- cut(df[, predmort],
                breaks = quantile(df[, predmort], probs = seq(0, 1, 0.20)), include.lowest = TRUE)
      
      df$risk_q5 <- factor(q5, levels = levels(q5), labels = c('q1', 'q2', 'q3', 'q4', "q5"))
      
      exp_risk <- tapply(df[, predmort], df[, "risk_q5"], mean)
      
      surv_fit <- survfit(Surv(df[, time], df[, status]) ~ df[, "risk_q5"])
      #obs_surv5 <- survfit(Surv(timeyr, delta) ~ q5_5yr_f, data = df)
      obs_risk <- 1 - summary(surv_fit ,times = tmax)$surv
      obs_se <- summary(surv_fit ,times = tmax)$std.err
      
      res_obs[i, ] <- obs_risk
      res_se[i, ] <- obs_se
      res_exp[i, ] <- exp_risk
    }

    # Pool
    pool_obs_q1 <- pool_caliplot(res_obs[, 1], res_se[, 1], nimp = nimp)
    pool_obs_q2 <- pool_caliplot(res_obs[, 2], res_se[, 2], nimp = nimp)
    pool_obs_q3 <- pool_caliplot(res_obs[, 3], res_se[, 3], nimp = nimp)
    pool_obs_q4 <- pool_caliplot(res_obs[, 4], res_se[, 4], nimp = nimp)
    pool_obs_q5 <- pool_caliplot(res_obs[, 5], res_se[, 5], nimp = nimp)
    
    pool_exp <- colMeans(res_exp)
    
    caliplot_df <- cbind(rbind(pool_obs_q1, pool_obs_q2, 
                               pool_obs_q3, pool_obs_q4, 
                               pool_obs_q5), pool_exp) %>%
      as.data.frame(.)
    names(caliplot_df) <- c("obs.y", "obs.lower", "obs.upper", "exp.x")
    
    return(caliplot_df)
}
```


## Functions for pooling O/E ratios and AUC
pool_OEratio computes observed risk and its standard errors, as well as the mean of the expected risk to calculate O/E ratio, i.e. the ratio between observed and expected risks. In addition, it also returns the number of events during different follow-up times.
```{r pooling function for OEratio}
pool_OEratio <- function(data, time, status, impvar, nimp, predmort, tmax) {
    res <- matrix(NA, nimp, 4) 
    # a results matrix for O/E ratio, se of O/E ratio, number of patients and number of events

    for (i in 1:nimp) {

      df <- data[data[impvar] == i, ]
      nwomen <- dim(df)[1]
      
      surv_fit <- survfit(Surv(df[, time], df[, status]) ~ 1)
      
      obs_risk <- 1 - summary(surv_fit, times = tmax)$surv
      obs_se <- summary(surv_fit, times = tmax)$std.err
      nevent <- summary(surv_fit, times = tmax)$n.event
      
      exp_risk <- mean(df[, predmort])
      
      # O/E ratio
      OE_stats <- obs_risk / exp_risk
      se_OE_stats <- obs_se / exp_risk

      res[i, ] <- c(OE_stats, se_OE_stats, nwomen, nevent)
    }

    # Pool
    pool_OE <- pool_estimate(res[, 1], res[, 2], nimp = nimp)
    
    nwomen <- round(median(res[, 3]), 0)
    nevent <- median(res[, 4])
    

    return(c(nwomen, nevent, pool_OE))
}
```

pool_AUC computes AUC and its 95% confidence interval. This function contains timeROC from timeROC pacakge.
```{r pooling function for AUC}
pool_AUC <- function(data, time, status, impvar, nimp, predmort, tmax) {
    res <- matrix(NA, nimp, 2) 

    for (i in 1:nimp) {

      df <- data[data[impvar] == i, ]
      
      tAUC <- timeROC(T = df[, time], delta = df[, status],
                  marker = df[, predmort], weighting="marginal", 
                  times = tmax, cause = 1, ROC = F, iid = TRUE)
      
      res[i, ] <- c(tAUC$AUC[2], tAUC$inference$vect_sd_1[2])
    }

    # Pool
    pool_AUCres <- pool_estimate(res[, 1], res[, 2], nimp = nimp)
  
    return(pool_AUCres)
}


```



# Performance of PREDICT v2.2
## Calibration in the large
```{r v22 OE ratio}
#10-year
OE10_v22_all <- pool_OEratio(df_v22, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_ERpos <- pool_OEratio(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_ERneg <- pool_OEratio(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_HRposHER2neg <- pool_OEratio(df_v22_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_HRposHER2pos <- pool_OEratio(df_v22_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_HRnegHER2pos <- pool_OEratio(df_v22_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_HRnegHER2neg <- pool_OEratio(df_v22_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_HRnegHER2neg_gBRCA12m  <- pool_OEratio(df_v22_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v22_HRnegHER2neg_gBRCA12wt <- pool_OEratio(df_v22_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)

res_OE10_v22 <- matrix(c(OE10_v22_all, OE10_v22_ERpos, OE10_v22_ERneg, 
                         OE10_v22_HRposHER2neg, OE10_v22_HRposHER2pos, OE10_v22_HRnegHER2pos, OE10_v22_HRnegHER2neg,
                         OE10_v22_HRnegHER2neg_gBRCA12m, OE10_v22_HRnegHER2neg_gBRCA12wt),
                    nrow=9, byrow=T,
                    dimnames = list(c('All women','ER-positive','ER-negative',
                                      'HRposHER2neg', 'HRposHER2pos', 'HRnegHER2pos', 'HRnegHER2neg',
                                      'gBRCA12m', 'gBRCA12wt'),
                                (c('v22_No.women', 'v22_No.event', 'v22_O/E ratio'))))

res_OE10_v22

#5-year and-15 year
OE5_v22_all <- pool_OEratio(df_v22, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_ERpos <- pool_OEratio(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_ERneg <- pool_OEratio(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_HRposHER2neg <- pool_OEratio(df_v22_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_HRposHER2pos <- pool_OEratio(df_v22_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_HRnegHER2pos <- pool_OEratio(df_v22_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_HRnegHER2neg <- pool_OEratio(df_v22_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_HRnegHER2neg_gBRCA12m  <- pool_OEratio(df_v22_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v22_HRnegHER2neg_gBRCA12wt <- pool_OEratio(df_v22_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)


OE15_v22_all <- pool_OEratio(df_v22, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_ERpos <- pool_OEratio(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_ERneg <- pool_OEratio(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_HRposHER2neg <- pool_OEratio(df_v22_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_HRposHER2pos <- pool_OEratio(df_v22_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_HRnegHER2pos <- pool_OEratio(df_v22_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_HRnegHER2neg <- pool_OEratio(df_v22_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_HRnegHER2neg_gBRCA12m  <- pool_OEratio(df_v22_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v22_HRnegHER2neg_gBRCA12wt <- pool_OEratio(df_v22_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
```


## Calibration plot
We made a calibration plot of PREDICT v2.2 in all women, women with ER-positive tumors and women with ER-negative tumors.
```{r calibratio plot v22}
calplot10_dfv22 <- pool_caliplot_df(df_v22, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
calplot10_dfv22_ERpos <- pool_caliplot_df(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
calplot10_dfv22_ERneg <- pool_caliplot_df(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
```

```{r calibration all women, fig.width=12, fig.height=12}
caliplot_dfv22_all <-
  ggplot(calplot10_dfv22, aes(x = exp.x, y = obs.y)) +
  geom_point(color = "black", size = 3, shape = 15) +
  geom_line(color = "black", size = 1) +
  #abline! I used geomline instead of abline because I could specify the length of the abline
  geom_line(data = data.frame(x = c(0, 0.41), y = c(0, 0.41)), 
            aes(x, y), color = "black", linetype = "dashed") +
  geom_errorbar(aes(ymin = obs.lower, ymax = obs.upper),
                width = 0.02, color = "black", size = 0.7) +
  #geom_vline(xintercept = 0.12, color = "red", linetype = "dashed") + 
  xlim(0, 0.41) + ylim(0, 0.41) +
  xlab("Predicted 10-year all-cause mortality") +
  ylab("Observed 10-year all-cause mortality") +
  ggtitle("A. Calibration plot in all women") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 20), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 16),
        axis.line = element_line(color = "black", linewidth = 0.8),
        axis.text = element_text(size = 12)) 
caliplot_dfv22_all
```

```{r calibration ERpos, fig.width=12, fig.height=12}
caliplot_dfv22_ERpos <-
  ggplot(calplot10_dfv22_ERpos, aes(x = exp.x, y = obs.y)) +
  geom_point(color = "black", size = 3, shape = 15) +
  geom_line(color = "black", size = 1) +
  #abline! I used geomline instead of abline because I could specify the length of the abline
  geom_line(data = data.frame(x = c(0, 0.41), y = c(0, 0.41)), 
            aes(x, y), color = "black", linetype = "dashed") +
  geom_errorbar(aes(ymin = obs.lower, ymax = obs.upper),
                width = 0.02, color = "black", size = 0.7) +
  #geom_vline(xintercept = 0.12, color = "red", linetype = "dashed") + 
  xlim(0, 0.41) + ylim(0, 0.41) +
  xlab("Predicted 10-year all-cause mortality") +
  ylab("Observed 10-year all-cause mortality") +
  ggtitle("B. Calibration plot in women with ER-positive tumors") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 20), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 16),
        axis.line = element_line(color = "black", linewidth = 0.8),
        axis.text = element_text(size = 12)) 
caliplot_dfv22_ERpos
```

```{r calibration ERneg, fig.width=12, fig.height=12}
caliplot_dfv22_ERneg <-
  ggplot(calplot10_dfv22_ERneg, aes(x = exp.x, y = obs.y)) +
  geom_point(color = "black", size = 3, shape = 15) +
  geom_line(color = "black", size = 1) +
  #abline! I used geomline instead of abline because I could specify the length of the abline
  geom_line(data = data.frame(x = c(0, 0.41), y = c(0, 0.41)), 
            aes(x, y), color = "black", linetype = "dashed") +
  geom_errorbar(aes(ymin = obs.lower, ymax = obs.upper),
                width = 0.02, color = "black", size = 0.7) +
  #geom_vline(xintercept = 0.12, color = "red", linetype = "dashed") + 
  xlim(0, 0.41) + ylim(0, 0.41) +
  xlab("Predicted 10-year all-cause mortality") +
  ylab("Observed 10-year all-cause mortality") +
  ggtitle("C. Calibration plot in women with ER-negative tumors") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 20), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 16),
        axis.line = element_line(color = "black", linewidth = 0.8),
        axis.text = element_text(size = 12)) 
caliplot_dfv22_ERneg
```
```{r combine plots, include=FALSE}
# create a ggplot with white background
plotnull <- ggplot() +
  theme(panel.background = element_rect(fill = "white"))

Combined_caliplot10_dfv22 <-
  ggarrange(caliplot_dfv22_all, plotnull, caliplot_dfv22_ERpos, caliplot_dfv22_ERneg, 
          ncol = 2, nrow = 2,
          heights = 1, widths = 1)
ggsave("/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/3_Results and plots/Plots/Combined_caliplot10_dfv22.png", 
       plot = Combined_caliplot10_dfv22, width = 15, height = 15, dpi = 600)
```
## Discrimination
Discrimination ability was calculated using time-dependent area under the ROC curve.
```{r v22 ROC}
#10-year
AUC10_v22_all <- pool_AUC(df_v22, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_ERpos <- pool_AUC(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_ERneg <- pool_AUC(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_HRposHER2neg <- pool_AUC(df_v22_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_HRposHER2pos <- pool_AUC(df_v22_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_HRnegHER2pos <- pool_AUC(df_v22_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_HRnegHER2neg <- pool_AUC(df_v22_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_HRnegHER2neg_gBRCA12m  <- pool_AUC(df_v22_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v22_HRnegHER2neg_gBRCA12wt <- pool_AUC(df_v22_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)

res_AUC10_v22 <- matrix(c(AUC10_v22_all, AUC10_v22_ERpos, AUC10_v22_ERneg, 
                         AUC10_v22_HRposHER2neg, AUC10_v22_HRposHER2pos, AUC10_v22_HRnegHER2pos, AUC10_v22_HRnegHER2neg,
                         AUC10_v22_HRnegHER2neg_gBRCA12m, AUC10_v22_HRnegHER2neg_gBRCA12wt),
                    nrow=9, byrow=T,
                    dimnames = list(c('All women','ER-positive','ER-negative',
                                      'HRposHER2neg', 'HRposHER2pos', 'HRnegHER2pos', 'HRnegHER2neg',
                                      'gBRCA12m', 'gBRCA12wt'),
                                (c('v22_AUC'))))

res_AUC10_v22

#5-year and 15-year
AUC5_v22_all <- pool_AUC(df_v22, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_ERpos <- pool_AUC(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_ERneg <- pool_AUC(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_HRposHER2neg <- pool_AUC(df_v22_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_HRposHER2pos <- pool_AUC(df_v22_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_HRnegHER2pos <- pool_AUC(df_v22_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_HRnegHER2neg <- pool_AUC(df_v22_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_HRnegHER2neg_gBRCA12m  <- pool_AUC(df_v22_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v22_HRnegHER2neg_gBRCA12wt <- pool_AUC(df_v22_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)


AUC15_v22_all <- pool_AUC(df_v22, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_ERpos <- pool_AUC(df_v22_ERpos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_ERneg <- pool_AUC(df_v22_ERneg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_HRposHER2neg <- pool_AUC(df_v22_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_HRposHER2pos <- pool_AUC(df_v22_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_HRnegHER2pos <- pool_AUC(df_v22_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_HRnegHER2neg <- pool_AUC(df_v22_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_HRnegHER2neg_gBRCA12m  <- pool_AUC(df_v22_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v22_HRnegHER2neg_gBRCA12wt <- pool_AUC(df_v22_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
```

# Performance of PREDICT v2.3
## Calibration in the large
```{r v23 OE ratio}
#10-year
OE10_v23_all <- pool_OEratio(df_v23, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_ERpos <- pool_OEratio(df_v23_ERpos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_ERneg <- pool_OEratio(df_v23_ERneg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_HRposHER2neg <- pool_OEratio(df_v23_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_HRposHER2pos <- pool_OEratio(df_v23_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_HRnegHER2pos <- pool_OEratio(df_v23_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_HRnegHER2neg <- pool_OEratio(df_v23_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_HRnegHER2neg_gBRCA12m  <- pool_OEratio(df_v23_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
OE10_v23_HRnegHER2neg_gBRCA12wt <- pool_OEratio(df_v23_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)

res_OE10_v23 <- matrix(c(OE10_v23_all, OE10_v23_ERpos, OE10_v23_ERneg, 
                         OE10_v23_HRposHER2neg, OE10_v23_HRposHER2pos, OE10_v23_HRnegHER2pos, OE10_v23_HRnegHER2neg,
                         OE10_v23_HRnegHER2neg_gBRCA12m, OE10_v23_HRnegHER2neg_gBRCA12wt),
                    nrow=9, byrow=T,
                    dimnames = list(c('All women','ER-positive','ER-negative',
                                      'HRposHER2neg', 'HRposHER2pos', 'HRnegHER2pos', 'HRnegHER2neg',
                                      'gBRCA12m', 'gBRCA12wt'),
                                (c('v23_No.women', 'v23_No.event', 'v23_O/E ratio'))))

res_OE10_v23

#5-year and-15 year
OE5_v23_all <- pool_OEratio(df_v23, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_ERpos <- pool_OEratio(df_v23_ERpos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_ERneg <- pool_OEratio(df_v23_ERneg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_HRposHER2neg <- pool_OEratio(df_v23_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_HRposHER2pos <- pool_OEratio(df_v23_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_HRnegHER2pos <- pool_OEratio(df_v23_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_HRnegHER2neg <- pool_OEratio(df_v23_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_HRnegHER2neg_gBRCA12m  <- pool_OEratio(df_v23_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
OE5_v23_HRnegHER2neg_gBRCA12wt <- pool_OEratio(df_v23_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)


OE15_v23_all <- pool_OEratio(df_v23, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_ERpos <- pool_OEratio(df_v23_ERpos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_ERneg <- pool_OEratio(df_v23_ERneg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_HRposHER2neg <- pool_OEratio(df_v23_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_HRposHER2pos <- pool_OEratio(df_v23_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_HRnegHER2pos <- pool_OEratio(df_v23_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_HRnegHER2neg <- pool_OEratio(df_v23_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_HRnegHER2neg_gBRCA12m  <- pool_OEratio(df_v23_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
OE15_v23_HRnegHER2neg_gBRCA12wt <- pool_OEratio(df_v23_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
```
## Discrimination
Discrimination ability was calculated using time-dependent area under the ROC curve.
```{r v23 ROC}
#10-year
AUC10_v23_all <- pool_AUC(df_v23, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_ERpos <- pool_AUC(df_v23_ERpos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_ERneg <- pool_AUC(df_v23_ERneg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_HRposHER2neg <- pool_AUC(df_v23_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_HRposHER2pos <- pool_AUC(df_v23_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_HRnegHER2pos <- pool_AUC(df_v23_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_HRnegHER2neg <- pool_AUC(df_v23_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_HRnegHER2neg_gBRCA12m  <- pool_AUC(df_v23_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)
AUC10_v23_HRnegHER2neg_gBRCA12wt <- pool_AUC(df_v23_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all10", tmax = 10)

res_AUC10_v23 <- matrix(c(AUC10_v23_all, AUC10_v23_ERpos, AUC10_v23_ERneg, 
                         AUC10_v23_HRposHER2neg, AUC10_v23_HRposHER2pos, AUC10_v23_HRnegHER2pos, AUC10_v23_HRnegHER2neg,
                         AUC10_v23_HRnegHER2neg_gBRCA12m, AUC10_v23_HRnegHER2neg_gBRCA12wt),
                    nrow=9, byrow=T,
                    dimnames = list(c('All women','ER-positive','ER-negative',
                                      'HRposHER2neg', 'HRposHER2pos', 'HRnegHER2pos', 'HRnegHER2neg',
                                      'gBRCA12m', 'gBRCA12wt'),
                                (c('v23_AUC'))))

res_AUC10_v23

#5-year and 15-year
AUC5_v23_all <- pool_AUC(df_v23, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_ERpos <- pool_AUC(df_v23_ERpos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_ERneg <- pool_AUC(df_v23_ERneg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_HRposHER2neg <- pool_AUC(df_v23_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_HRposHER2pos <- pool_AUC(df_v23_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_HRnegHER2pos <- pool_AUC(df_v23_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_HRnegHER2neg <- pool_AUC(df_v23_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_HRnegHER2neg_gBRCA12m  <- pool_AUC(df_v23_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)
AUC5_v23_HRnegHER2neg_gBRCA12wt <- pool_AUC(df_v23_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all5", tmax = 5)


AUC15_v23_all <- pool_AUC(df_v23, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_ERpos <- pool_AUC(df_v23_ERpos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_ERneg <- pool_AUC(df_v23_ERneg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_HRposHER2neg <- pool_AUC(df_v23_HRposHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_HRposHER2pos <- pool_AUC(df_v23_HRposHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_HRnegHER2pos <- pool_AUC(df_v23_HRnegHER2pos, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_HRnegHER2neg <- pool_AUC(df_v23_HRnegHER2neg, "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_HRnegHER2neg_gBRCA12m  <- pool_AUC(df_v23_HRnegHER2neg_gBRCA12m , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
AUC15_v23_HRnegHER2neg_gBRCA12wt <- pool_AUC(df_v23_HRnegHER2neg_gBRCA12wt , "timeyr", "delta", ".imp", 50, "all15", tmax = 15)
```


## Results summary
### Table 3
```{r res_table3}
res_10 <- cbind(res_OE10_v22, res_AUC10_v22, res_OE10_v23, res_AUC10_v23) %>%
  as_tibble() 
res_10
```
### Suapplmentary Table 1
```{r res_sTable1}
s_res_v22 <- matrix(c(OE5_v22_all, AUC5_v22_all, 
                    OE15_v22_all, AUC15_v22_all, 
                    OE5_v22_ERpos, AUC5_v22_ERpos, 
                    OE15_v22_ERpos, AUC15_v22_ERpos,
                    OE5_v22_ERneg, AUC5_v22_ERneg, 
                    OE15_v22_ERneg, AUC15_v22_ERneg,
                    OE5_v22_HRposHER2neg, AUC5_v22_HRposHER2neg, 
                    OE15_v22_HRposHER2neg, AUC15_v22_HRposHER2neg, 
                    OE5_v22_HRposHER2pos, AUC5_v22_HRposHER2pos, 
                    OE15_v22_HRposHER2pos, AUC15_v22_HRposHER2pos, 
                    OE5_v22_HRnegHER2pos, AUC5_v22_HRnegHER2pos, 
                    OE15_v22_HRnegHER2pos, AUC15_v22_HRnegHER2pos,
                    OE5_v22_HRnegHER2neg, AUC5_v22_HRnegHER2neg, 
                    OE15_v22_HRnegHER2neg, AUC15_v22_HRnegHER2neg,
                    OE5_v22_HRnegHER2neg_gBRCA12m, AUC5_v22_HRnegHER2neg_gBRCA12m, 
                    OE15_v22_HRnegHER2neg_gBRCA12m, AUC15_v22_HRnegHER2neg_gBRCA12m, 
                    OE5_v22_HRnegHER2neg_gBRCA12wt, AUC5_v22_HRnegHER2neg_gBRCA12wt, 
                    OE15_v22_HRnegHER2neg_gBRCA12wt, AUC15_v22_HRnegHER2neg_gBRCA12wt),
                    nrow=9, byrow=T,
                    dimnames = list(c('All women','ER-positive','ER-negative',
                                      'HRposHER2neg', 'HRposHER2pos', 'HRnegHER2pos', 'HRnegHER2neg',
                                      'gBRCA12m', 'gBRCA12wt'),
                                (c('No.women5', 'No.event5', 'O/E ratio5', 'AUC5',
                                   'No.women15', 'No.event15', 'O/E ratio15', 'AUC15'))))

s_res_v22
```

### Suapplmentary Table 2
```{r res_sTable2}
s_res_v23 <- matrix(c(OE5_v23_all, AUC5_v23_all, 
                    OE15_v23_all, AUC15_v23_all, 
                    OE5_v23_ERpos, AUC5_v23_ERpos, 
                    OE15_v23_ERpos, AUC15_v23_ERpos,
                    OE5_v23_ERneg, AUC5_v23_ERneg, 
                    OE15_v23_ERneg, AUC15_v23_ERneg,
                    OE5_v23_HRposHER2neg, AUC5_v23_HRposHER2neg, 
                    OE15_v23_HRposHER2neg, AUC15_v23_HRposHER2neg, 
                    OE5_v23_HRposHER2pos, AUC5_v23_HRposHER2pos, 
                    OE15_v23_HRposHER2pos, AUC15_v23_HRposHER2pos, 
                    OE5_v23_HRnegHER2pos, AUC5_v23_HRnegHER2pos, 
                    OE15_v23_HRnegHER2pos, AUC15_v23_HRnegHER2pos,
                    OE5_v23_HRnegHER2neg, AUC5_v23_HRnegHER2neg, 
                    OE15_v23_HRnegHER2neg, AUC15_v23_HRnegHER2neg,
                    OE5_v23_HRnegHER2neg_gBRCA12m, AUC5_v23_HRnegHER2neg_gBRCA12m, 
                    OE15_v23_HRnegHER2neg_gBRCA12m, AUC15_v23_HRnegHER2neg_gBRCA12m, 
                    OE5_v23_HRnegHER2neg_gBRCA12wt, AUC5_v23_HRnegHER2neg_gBRCA12wt, 
                    OE15_v23_HRnegHER2neg_gBRCA12wt, AUC15_v23_HRnegHER2neg_gBRCA12wt),
                    nrow=9, byrow=T,
                    dimnames = list(c('All women','ER-positive','ER-negative',
                                      'HRposHER2neg', 'HRposHER2pos', 'HRnegHER2pos', 'HRnegHER2neg',
                                      'gBRCA12m', 'gBRCA12wt'),
                                (c('No.women5', 'No.event5', 'O/E ratio5', 'AUC5',
                                   'No.women15', 'No.event15', 'O/E ratio15', 'AUC15'))))

s_res_v23
```

## Session information
```{r sinfo}
sessionInfo()
```

