---
title: "PREDICT_PARADIGM: Clinical Utility of PREDICT v2.2 and v2.3"
author: "Yuwei Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
  toc_depth: 3
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# READ ME
The R script is designed for our manuscript entitled the external validation and clinical utility assessment of PREDICT breast cancer prognostic model in young women with node-negative breast cancer. This R script assessed the potential clinical utility of PREDICT v2.2 and PREDICT v2.3. Using the script, we generated results for Table 4 and Supplementary Table 5.

# Load the libraries
```{r load the libraries, message=FALSE, warning=FALSE}
library(readxl)
library(readr)
library(dplyr)
library(plyr)
library(rms)
library(ggplot2)
library(survminer)
library(survival)
```

# Load the imputed datasets
```{r load imputed data, include=F}
dataimpute_v22 <- read.csv(file='/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/dataimpute_v22_July.csv',stringsAsFactors=F,header=T,sep = ",") %>%
  dplyr::select(-X)

dataimpute_v23 <- read.csv(file='/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/dataimpute_v23_July.csv',stringsAsFactors=F,header=T,sep = ",") %>%
  dplyr::select(-X)
```

```{r datasets for model v22}
## women with ER-positive tumors
df_v22_ERpos <- dataimpute_v22 %>%
        filter(ER.num > 0 & T_LEEFT >= 25)

## women with ER-negative tumors
df_v22_ERneg <- dataimpute_v22 %>%
        filter(ER.num == 0 & T_LEEFT >= 25)
```

```{r datasets for model v23}
## women with ER-positive tumors
df_v23_ERpos <- dataimpute_v23 %>%
        filter(ER.num > 0 & T_LEEFT >= 25)

## women with ER-negative tumors
df_v23_ERneg <- dataimpute_v23 %>%
        filter(ER.num == 0 & T_LEEFT >= 25)
```


# Net benefit function
We made a function for pooling net benefits at predefined thresholds in each imputed dataset.
```{r model for net benefit, warning=FALSE}
NBFun <- function (model, imputedata, nimp, impvar, thresholdrange, predrisk, time, status, tmax, totalnumber){

  if(model == "Chemotherapy-to-all") {
    
    pool_res_allthreshold <- matrix(NA, length(thresholdrange), 3)
    
        for (q in thresholdrange){
          
          weight <- q/(1 - q)
          
          res_in_each_impute <- matrix(NA, nimp, 3)
          
          for (i in 1:nimp){
      
            df <- imputedata[imputedata[impvar] == i,]
            
            df_predpos <- df
            #all are "positive" since this is treatment for all
            surv_fit_pos <- survfit(Surv(df_predpos[, time], df_predpos[, status]) ~ 1)
            TP <- 1 - summary(surv_fit_pos, times = tmax)$surv
            TPrate <- TP * nrow(df_predpos)/nrow(df_predpos)
                            
            FP <- summary(surv_fit_pos, times = tmax)$surv
            FPrate <- FP * nrow(df_predpos)/nrow(df_predpos)
            
            NB <- TPrate - weight * FPrate
            
            res_in_each_impute[i, ] <- c(TPrate, FPrate, NB)
          
          }
          
      pool_res <- round(colMeans(res_in_each_impute) * totalnumber, 0)
      
      q_loc <- which(thresholdrange == q)
      pool_res_allthreshold[q_loc , ] <- pool_res
        }
      
    res <- cbind(c(thresholdrange), pool_res_allthreshold)
    
    colnames(res) <- c(paste0(model, "_threshold"), paste0(model, "_TP"), 
                       paste0(model, "_FP"), paste0(model, "_NB"))
    
    return(res)

  }
  
  if(model != "Chemotherapy-to-all") {
        
    pool_res_allthreshold <- matrix(NA, length(thresholdrange), 5)  
    
        for (q in thresholdrange){
          
          weight <- q/(1 - q)
          
          res_in_each_impute <- matrix(NA, nimp, 5) 
          
          for (i in 1:nimp){
      
          df <- imputedata[imputedata[impvar] == i,]
          
          df_predpos <- df[df[predrisk] >= q, ]
          
          if(dim(df_predpos)[1] > 0){
            
              surv_fit_pos <- survfit(Surv(df_predpos[, time], df_predpos[, status]) ~ 1)
              TP <- 1 - summary(surv_fit_pos, times = tmax)$surv
              TPrate <- TP * nrow(df_predpos)/nrow(df)
                              
              FP <- summary(surv_fit_pos, times = tmax)$surv
              FPrate <- FP * nrow(df_predpos)/nrow(df)
          
          } else{
            
            TPrate = 0
            FPrate = 0
          }
          
          df_predneg <- df[df[predrisk] < q, ]
          
          if(dim(df_predneg)[1] > 0){
          
            surv_fit_neg <- survfit(Surv(df_predneg[, time], df_predneg[, status]) ~ 1)
            FN <- 1 - summary(surv_fit_neg, times = tmax)$surv
            FNrate <- FN * nrow(df_predneg)/nrow(df)
    
            TN <- summary(surv_fit_neg, times = tmax)$surv
            TNrate <- TN * nrow(df_predneg)/nrow(df)
          } else{
            
            TNrate = 0
            FNrate = 0        
          }   
          
          NB <- TPrate - weight * FPrate
          
          res_in_each_impute[i, ] <- c(TPrate, FPrate, NB, FNrate, TNrate)
          
          }
      pool_res <- round(colMeans(res_in_each_impute) * totalnumber, 0)
      
      q_loc <- which(thresholdrange == q)
      pool_res_allthreshold[q_loc , ] <- pool_res
        }
    
    res <- cbind(c(thresholdrange), pool_res_allthreshold)
    
    colnames(res) <- c(paste0(model, "_threshold"), paste0(model, "_TP"), paste0(model, "_FP"), 
                       paste0(model, "_NB"), paste0(model, "_FN"), paste0(model, "_TN"))
    
    return(res)

  }
}
```

# Net benefits of PREDICT v2.2, v2.3, and chemotherapy-to-all

## Threshold
```{r}
threshold <- seq(0.05, 0.15, by = 0.005)
```

## PREDICT v2.2
```{r v22}
res.ERpos_v22 <- NBFun(model = "PREDICTv22", imputedata = df_v22_ERpos, nimp = 50, 
                impvar = ".imp", thresholdrange = threshold, predrisk = "all10", 
                time = "timeyr", status = "delta", tmax = 10, totalnumber = 1000)
res.ERpos_v22

res.ERneg_v22 <- NBFun(model = "PREDICTv22", imputedata = df_v22_ERneg, nimp = 50, 
                impvar = ".imp", thresholdrange = threshold, predrisk = "all10", 
                time = "timeyr", status = "delta", tmax = 10, totalnumber = 1000)
res.ERneg_v22
```

## PREDICT v2.3
```{r v23}
res.ERpos_v23 <- NBFun(model = "PREDICTv23", imputedata = df_v23_ERpos, nimp = 50, 
                impvar = ".imp", thresholdrange = threshold, predrisk = "all10", 
                time = "timeyr", status = "delta", tmax = 10, totalnumber = 1000)
res.ERpos_v23

res.ERneg_v23 <- NBFun(model = "PREDICTv23", imputedata = df_v23_ERneg, nimp = 50, 
                impvar = ".imp", thresholdrange = threshold, predrisk = "all10", 
                time = "timeyr", status = "delta", tmax = 10, totalnumber = 1000)
res.ERneg_v23
```

## Chemotherapy-to-all
To compute net benefits for chemotherapy-to-all strategy, it doesn't matter if we use df_v22 or df_v23. This is because chemotherapy-to-all classify all patients as high-risk, and the only difference between df_v22 and df_v23 are the predicted risks.
```{r chemotoall}
res.ERpos_chemotoall <- NBFun(model = "Chemotherapy-to-all", imputedata = df_v22_ERpos, nimp = 50, 
                impvar = ".imp", thresholdrange = threshold, predrisk = "all10", 
                time = "timeyr", status = "delta", tmax = 10, totalnumber = 1000)
res.ERpos_chemotoall

res.ERneg_chemotoall <- NBFun(model = "Chemotherapy-to-all", imputedata = df_v22_ERneg, nimp = 50, 
                impvar = ".imp", thresholdrange = threshold, predrisk = "all10", 
                time = "timeyr", status = "delta", tmax = 10, totalnumber = 1000)
res.ERneg_chemotoall
```

# Results
## Results for Supplementary Table 5
```{r results ST5}
res.ERpos <- cbind(res.ERpos_v22, res.ERpos_v23, res.ERpos_chemotoall)
res.ERneg <- cbind(res.ERneg_v22, res.ERneg_v23, res.ERneg_chemotoall)

stable5.ERpos <- res.ERpos[seq(1, 21, by = 2), c(1:6, 8:12, 14:16)] %>% as_tibble(.)
stable5.ERpos$diff_v22 <- stable5.ERpos$PREDICTv22_NB - stable5.ERpos$`Chemotherapy-to-all_NB`
stable5.ERpos$diff_v23 <- stable5.ERpos$PREDICTv23_NB - stable5.ERpos$`Chemotherapy-to-all_NB`

stable5.ERneg <- res.ERneg[seq(1, 21, by = 2), c(1:6, 8:12, 14:16)] %>% as_tibble(.)
stable5.ERneg$diff_v22 <- stable5.ERneg$PREDICTv22_NB - stable5.ERneg$`Chemotherapy-to-all_NB`
stable5.ERneg$diff_v23 <- stable5.ERneg$PREDICTv23_NB - stable5.ERneg$`Chemotherapy-to-all_NB`

stable5.ERpos
stable5.ERneg
```

## Make the decision curve plots

```{r DCA plot dataset}
df.colname <- c("threshold", "treatall_NB", "PREDICTv22_NB", "PREDICTv23_NB")

df.plot.pos <- cbind(res.ERpos_chemotoall[, c("Chemotherapy-to-all_threshold", "Chemotherapy-to-all_NB")], 
                    res.ERpos_v22[, "PREDICTv22_NB"], res.ERpos_v23[, "PREDICTv23_NB"]) %>%
  as_tibble()

names(df.plot.pos) <- df.colname

df.plot.neg <- cbind(res.ERneg_chemotoall[, c("Chemotherapy-to-all_threshold", "Chemotherapy-to-all_NB")], 
                    res.ERneg_v22[, "PREDICTv22_NB"], res.ERneg_v23[, "PREDICTv23_NB"])%>%
  as_tibble()

names(df.plot.neg) <- df.colname
```

```{r plot ERpos, fig.height=10, fig.width=10}
plot.ERpos <- 
  ggplot() +
  geom_line(data = df.plot.pos,
            aes(x = threshold, y = PREDICTv22_NB, color = "PREDICT v2.2"), linewidth = 1) +
  geom_line(data = df.plot.pos, 
            aes(x = threshold, y = PREDICTv23_NB, color = "PREDICT v2.3"), linewidth = 1) +
  geom_line(data = df.plot.pos, 
            aes(x = threshold, y = treatall_NB, color = "Chemotherapy-to-all"), linewidth = 1) +
  geom_vline(xintercept = 0.12, color = "black", linetype = "dashed") +
  # geom_vline(xintercept = 0.10, color = "black", linetype = "dotted") +
  # geom_vline(xintercept = 0.15, color = "black", linetype = "dotted") +
  xlim(0.05, 0.15) + ylim(65, 215) +
  scale_x_continuous(breaks = seq(0.05, 0.15, by = 0.01)) +
  scale_y_continuous(breaks = seq(65, 215, by = 15)) +
  xlab("Risk Classification Threshold") +
  ylab("Net benefit per 1000") +
  ggtitle("A. Decision Curve Analysis in Women with ER-positive tumors") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 25), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 22),
        axis.text = element_text(size = 18),
        axis.line = element_line(color = "black", linewidth = 1),
        legend.position = c(0.82, 0.85),
        legend.title = element_blank(),
        legend.key.size = unit(1.5, "cm"),
        legend.text = element_text(size = 22)) +
  scale_color_manual(values = c("PREDICT v2.2" = "#B2182B", 
                                "PREDICT v2.3" = "#4393C3",
                                "Chemotherapy-to-all" = "darkgray"))
plot.ERpos
```

```{r plot ER-neg, fig.height=10, fig.width=10}
plot.ERneg <- 
  ggplot() +
  geom_line(data = df.plot.neg,
            aes(x = threshold, y = PREDICTv22_NB, color = "PREDICT v2.2"), linewidth = 1) +
  geom_line(data = df.plot.neg, 
            aes(x = threshold, y = PREDICTv23_NB, color = "PREDICT v2.3"), linewidth = 1) +
  geom_line(data = df.plot.neg, 
            aes(x = threshold, y = treatall_NB, color = "Chemotherapy-to-all"), linewidth = 1) +
  geom_vline(xintercept = 0.08, color = "black", linetype = "dashed") +
  #geom_vline(xintercept = 0.07, color = "black", linetype = "dotted") +
  #geom_vline(xintercept = 0.12, color = "black", linetype = "dotted") +
  xlim(0.05, 0.15) + ylim(65, 215) +
  scale_x_continuous(breaks = seq(0.05, 0.15, by = 0.01)) +
  scale_y_continuous(breaks = seq(65, 215, by = 15)) +
  xlab("Risk Classification Threshold") +
  ylab("Net benefit per 1000") +
  ggtitle("B. Decision Curve Analysis in Women with ER-negative tumors") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 25), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 22),
        axis.text = element_text(size = 18),
        axis.line = element_line(color = "black", linewidth = 1),
        legend.position = c(0.82, 0.85),
        legend.title = element_blank(),
        legend.key.size = unit(1.5, "cm"),
        legend.text = element_text(size = 22)) +
  scale_color_manual(values = c("PREDICT v2.2" = "#B2182B", 
                                "PREDICT v2.3" = "#4393C3",
                                "Chemotherapy-to-all" = "darkgray"))
plot.ERneg
```

```{r plot combined}
setwd("/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/3_Results and plots/Plots")
png("DCA_combined.png", width=2000, height=1000)
plot.ERneg
gridExtra::grid.arrange(plot.ERpos, plot.ERneg, ncol = 2)

```
