---
title: "PREDICT_PARADIGM: 4. complete cases"
author: "Yuwei Wang"
date:  "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
  toc_depth: 3  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# READ ME
The R script is designed for our manuscript entitled the external validation and clinical utility assessment of PREDICT breast cancer prognostic model in young women with node-negative breast cancer. This R script assessed the model validity of PREDICT v2.2 and PREDICT v2.3 in women with complete information on tumor ER status, size, grade, . Using the script, we generated results for Supplementary Table 3 - 4.

# Load the libraries
```{r load the libraries, message=FALSE}
library("readxl")
library("readr")
library("dplyr")
library("plyr")
library("rms")
library("ggplot2")
library("timeROC")
library("stringr")
library("tidyverse")

```

# Load the complete dataset
```{r load the PARADIGM data set, include=FALSE}
TNBC <- read.csv(file='/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/TNBC485_BRCA1statusupdate_in_August_2022.csv',stringsAsFactors=F,header=T,sep = ",") %>%
  filter(T_LEEFT >= 25 & ER.num == 0 & PR.num == 0) 
gBRCAmnr <- TNBC$palga_nr[which(TNBC$gBRCA1 == 1 | TNBC$BRCA2 == "gBRCA2m")]
gBRCAwtnr<- TNBC$palga_nr[which(TNBC$gBRCA1 == 0 & TNBC$BRCA2 %in% c("sBRCA2m", "BRCA2wt"))]

PRD <- read.csv("/DATA/users/yu.wang/8_PREDICT/PREDICT_PARADIGM_project/1_Data/nbcp1.csv",stringsAsFactors=F,header=T,sep = ",") %>%
    dplyr::select(-X) %>%
    mutate(incyear = incdat,
           age.start = T_LEEFT,
           er = factor(ifelse(ER.num == 0, 0, 
                          ifelse(ER.num > 0, 1, NA))),
           pr = factor(ifelse(PR.num == 0, 0, 
                            ifelse(PR.num > 0, 1, NA))),
           her2 = factor(ifelse(HER2.cat == "Negative", 0,
                         ifelse(HER2.cat == "Positive", 1, NA))),
           size = tumorsize,
           grade = factor(ifelse(Grade == "Grade 1", 1,
                          ifelse(Grade == "Grade 2", 2,
                                 ifelse(Grade == "Grade 3", 3, NA)))),
           ki67 = 9,
           screen = 0,
           nodes = 0,
           gBRCA12 = ifelse(palga_nr %in% gBRCAmnr, 1,
                            ifelse(palga_nr %in% gBRCAwtnr, 0, NA)),
           
           Radiotherapy = factor(ifelse(is.na(Radiotherapy), "No", Radiotherapy)),
           Surgery.cat3 = factor(ifelse(str_detect(Surgery, "lumpectomy"), "Lumpectomy",
                                 ifelse(str_detect(Surgery, "mastectomy"), "Mastectomy", "Surgery NNO"))),
           Surgery = factor(Surgery),
           
           delta = ifelse(fustat.final == "overleden 1", 1, 0),
           timeyr = VITFUP.final/365.25,
           intrsub = ifelse((er == 1| pr == 1) & her2 == 0, "HRposHER2neg",
                              ifelse((er == 1| pr == 1) & her2 == 1, "HRposHER2pos",
                                     ifelse(er == 0 & pr == 0 & her2 == 1, "HRnegHER2pos",
                                            ifelse(er == 0 & pr == 0 & her2 == 0, "HRnegHER2neg", NA))))) %>%
  separate(incyear, c("incyear", "incmonth"), "-", extra = "merge", fill = "right")

```

```{r load the complete dataset}
comp_df <- PRD %>%
  filter(age.start >= 25) %>%
  dplyr::select(palga_nr, age.start, screen, size, grade, nodes, er, pr, her2, ki67, delta, timeyr, gBRCA12, intrsub) %>%
  .[complete.cases(.$palga_nr, .$age.start, .$screen, .$size, .$grade, .$nodes, .$er, .$pr, .$her2, .$ki67, .$delta, .$timeyr),]

```

# Load functions
```{r functions for validation}
validity_Fun <- function(data, time, status, predmort, tmax) {
      nwomen <- dim(data)[1]
      ntime <- length(tmax)
      res_validity <- c(nwomen)
      for (i in 1:ntime){
        
        surv_fit <- survfit(Surv(data[, time], data[, status]) ~ 1)
        
        #calculate the OE ratio
        obs_risk <- 1 - summary(surv_fit, times = tmax[i])$surv
        obs_se <- summary(surv_fit, times = tmax[i])$std.err
        nevent <- summary(surv_fit, times = tmax[i])$n.event
        exp_risk <- mean(data[, predmort[i]])
        

        OE_stats <- obs_risk / exp_risk
        se_OE_stats <- obs_se / exp_risk
        lowerCI_OE_stats <- round(OE_stats - qnorm(1 - 0.05/2) * se_OE_stats, 2)
        upperCI_OE_stats <- round(OE_stats + qnorm(1 - 0.05/2) * se_OE_stats, 2)
        
        res_OE <- paste0(round(OE_stats, 2), " (", lowerCI_OE_stats, "-", upperCI_OE_stats, ")")
        
        
        #calculate the AUC
        AUC <- timeROC(T = data[, time], delta = data[, status],
                       marker = data[, predmort[i]], 
                       weighting = "marginal", 
                       times = tmax[i], cause = 1, ROC = F, iid = T)
        AUC_stats <- AUC$AUC[2]
        se_AUC_stats <- AUC$inference$vect_sd_1 [2]
        lowerCI_AUC_stats <- round(AUC_stats - qnorm(1 - 0.05/2) * se_AUC_stats, 2)
        upperCI_AUC_stats <- round(AUC_stats + qnorm(1 - 0.05/2) * se_AUC_stats, 2)
        
        res_AUC <- paste0(round(AUC_stats, 2), " (", lowerCI_AUC_stats, "-", upperCI_AUC_stats, ")")
        
        
        res_validity <- append(res_validity, c(nevent, res_OE, res_AUC))

      }
      return(res_validity)
}
```

```{r calidata function}
caliplot_df <- function(data, time, status, predmort, tmax) {

    q5 <- cut(data[, predmort],
              breaks = quantile(data[, predmort], probs = seq(0, 1, 0.20)), include.lowest = TRUE)
    
    data$risk_q5 <- factor(q5, levels = levels(q5), labels = c('q1', 'q2', 'q3', 'q4', "q5"))
    
    exp_risk <- tapply(data[, predmort], data[, "risk_q5"], mean)
    
    surv_fit <- survfit(Surv(data[, time], data[, status]) ~ data[, "risk_q5"])
    
    obs_risk <- 1 - summary(surv_fit ,times = tmax)$surv
    obs_se <- summary(surv_fit ,times = tmax)$std.err
    
    obs_lower <- obs_risk - qnorm(1 - 0.05/2) * obs_se
    obs_upper <- obs_risk + qnorm(1 - 0.05/2) * obs_se
      
    
    caliplot_data <- cbind(obs_risk, obs_lower, obs_upper, exp_risk) %>%
      as.data.frame(.)
    names(caliplot_data) <- c("obs.y", "obs.lower", "obs.upper", "exp.x")
    
    return(caliplot_data)
}

```

# Run the script
## PREDICT V2.2
Note that PREDICTv2.2 published on https://github.com/gmp26/v22/blob/master/R/v22_model.R was slightly different.
  1. PREDICTv2.2 allows missing values in grade (if missing, impute missingness with 2.13) and screening (if missing, impute missingness with 0.204). In my script, grade is not allowed to be missing, the same as the PREDICT published on their website. For screening, we expect very few women in our cohort who would undergo screening, as they were all under 40 years at diagnosis. Screening for BRCA1/2 carriers was not yet inplemented before 2000.
  2. PREDICTv2.2 on the github added several codes about radiotherapy. However, these codes are currently disabled thus not being used right now.
  3. PREDICTv2.2 on the github included systemic treatment. I deleted these part because all women in our cohort were adjuvant treatment-naive.
```{r Run the datasest in the PREDICTv22 script}
for (n in 1:nrow(comp_df)) {
  # Input case characteristics
  age.start  <- comp_df$age.start[n]
  screen     <- comp_df$screen[n]     # Clinically detected = 0, screen detected = 1
  size       <- comp_df$size[n]   # Tumour size mm
  grade      <- comp_df$grade[n]     # Tumour grade
  nodes      <- comp_df$nodes[n]     # Number positive nodes. Nodal micrometastases = 0.5
  er         <- comp_df$er[n]     # ER+ = 1, ER- = 0
  her2       <- comp_df$her2[n]     # HER2+ = 1, HER2- = 0, missing = 9
  ki67       <- comp_df$ki67[n]     # KI67+ = 1, KI67- = 0, missing = 9

  # Grade variable for ER neg
  grade.val <- ifelse(er==1, grade, ifelse(grade==1, 0, 1))

  # Generate the coefficients
  age.mfp.1   <- ifelse(er==1, (age.start/10)^-2-.0287449295, age.start-56.3254902)
  age.beta.1  <- ifelse(er==1, 34.53642, 0.0089827)
  age.mfp.2   <- ifelse(er==1, (age.start/10)^-2*log(age.start/10)-.0510121013, 0)
  age.beta.2  <- ifelse(er==1, -34.20342, 0)
  size.mfp    <- ifelse(er==1, log(size/100)+1.545233938, (size/100)^.5-.5090456276)
  size.beta   <- ifelse(er==1, 0.7530729, 2.093446)
  nodes.mfp   <- ifelse(er==1,log((nodes+1)/10)+1.387566896,
                        log((nodes+1)/10)+1.086916249)
  nodes.beta  <- ifelse(er==1, 0.7060723, .6260541)
  grade.beta  <- ifelse(er==1, 0.746655, 1.129091)
  screen.beta <- ifelse(er==1, -0.22763366, 0)
  her2.beta   <- ifelse(her2==1, 0.2413,
                        ifelse(her2==0, -0.0762, 0))
  ki67.beta   <- ifelse(ki67==1 & er==1, 0.14904,
                        ifelse(ki67==0 & er==1, -0.1133, 0))

  # Calculate the other and breast mortality indicies

  # Other mortality prognostic index (mi)
  mi <- 0.0698252*((age.start/10)^2-34.23391957)

  # Breast cancer mortality prognostic index (pi)
  pi <- age.beta.1*age.mfp.1 + age.beta.2*age.mfp.2 + size.beta*size.mfp +
    nodes.beta*nodes.mfp + grade.beta*grade.val + screen.beta*screen +
    her2.beta + ki67.beta

  # Non breast cancer mortality
  # Generate cumulative baseline other mortality

  time<-c(5,10,15)

  base.m.cum.oth <- exp(-6.052919 + (1.079863*log(time)) + (.3255321*time^.5))


  # Generate cumulative survival non-breast mortality
  #YW: cumulative non-breast cancer survival
  s.cum.oth <- exp(-exp(mi)*base.m.cum.oth)

  # Generate annual survival from cumulative survival
  #YW: cumulative non-breast cancer mortality
  m.cum.oth <- 1 - s.cum.oth

  # Breast cancer specific mortality
  # Generate cumulative baseline breast mortality
  if (er==1) {
    base.m.cum.br <- exp(0.7424402 - 7.527762/time^.5 - 1.812513*log(time)/time^.5)
  } else { base.m.cum.br <- exp(-1.156036 + 0.4707332/time^2 - 3.51355/time)
  }
  # Calculate the cumulative breast cancer survival
  s.cum.br <- exp(-exp(pi)*base.m.cum.br)
  m.cum.br <- 1 - s.cum.br

  # All cause mortality
  m.cum.all <- 1 - s.cum.oth*s.cum.br
  s.cum.all <- 100-100*m.cum.all

  # Proportion of all cause mortality that is breast cancer
  prop.br <- m.cum.br/(m.cum.br+m.cum.oth)
  prop.oth <- m.cum.oth/(m.cum.br+m.cum.oth)
  # Predicted cumulative breast specific mortality
  pred.m.br    <- prop.br*m.cum.all
  # Predicted cumulative non-breast cancer mortality
  pred.m.oth <- prop.oth*m.cum.all
  # Predicted cumulative all-cause mortality
  pred.all <- pred.m.br + pred.m.oth #pred.all=m.cum.all

  # Predicted all cause mortality
  comp_df$v22_all5[n] <- pred.all[1]
  comp_df$v22_all10[n] <- pred.all[2]
  comp_df$v22_all15[n] <- pred.all[3]


  #prognostic index, this is not related to time
  comp_df$v22_pi[n] <- pi #larger pi is associated with higher risks of events

}
```
## PREDICT V2.3
PREDICT v2.3 was not yet published. The information on the predictor PR was derived from the manuscript from Grootes et al. 2022.
```{r Run the datasest in the PREDICTv223 script}
for (n in 1:nrow(comp_df)) {
  # Input case characteristics
  age.start  <- comp_df$age.start[n]
  screen     <- comp_df$screen[n]     # Clinically detected = 0, screen detected = 1
  size       <- comp_df$size[n]   # Tumour size mm
  grade      <- comp_df$grade[n]     # Tumour grade
  nodes      <- comp_df$nodes[n]     # Number positive nodes. Nodal micrometastases = 0.5
  er         <- comp_df$er[n]     # ER+ = 1, ER- = 0
  her2       <- comp_df$her2[n]     # HER2+ = 1, HER2- = 0, missing = 9
  ki67       <- comp_df$ki67[n]     # KI67+ = 1, KI67- = 0, missing = 9
  pr       <- comp_df$pr[n]
  # Grade variable for ER neg
  grade.val <- ifelse(er==1, grade, ifelse(grade==1, 0, 1))

  # Generate the coefficients
  age.mfp.1   <- ifelse(er==1, (age.start/10)^-2-.0287449295, age.start-56.3254902)
  age.beta.1  <- ifelse(er==1, 34.53642, 0.0089827)
  age.mfp.2   <- ifelse(er==1, (age.start/10)^-2*log(age.start/10)-.0510121013, 0)
  age.beta.2  <- ifelse(er==1, -34.20342, 0)
  size.mfp    <- ifelse(er==1, log(size/100)+1.545233938, (size/100)^.5-.5090456276)
  size.beta   <- ifelse(er==1, 0.7530729, 2.093446)
  nodes.mfp   <- ifelse(er==1,log((nodes+1)/10)+1.387566896,
                        log((nodes+1)/10)+1.086916249)
  nodes.beta  <- ifelse(er==1, 0.7060723, .6260541)
  grade.beta  <- ifelse(er==1, 0.746655, 1.129091)
  screen.beta <- ifelse(er==1, -0.22763366, 0)
  her2.beta   <- ifelse(her2==1, 0.2413,
                        ifelse(her2==0, -0.0762, 0))
  ki67.beta   <- ifelse(ki67==1 & er==1, 0.14904,
                        ifelse(ki67==0 & er==1, -0.1133, 0))
  pr.beta     <- ifelse(pr == 1 & er == 1, -0.0618754,
                      ifelse(pr == 0 & er == 1, 0.2623643,
                             ifelse(pr == 1 & er == 0, -0.2231436, 0.0295588)))

  # Calculate the other and breast mortality indicies

  # Other mortality prognostic index (mi)
  mi <- 0.0698252*((age.start/10)^2-34.23391957)

  # Breast cancer mortality prognostic index (pi)
  pi <- age.beta.1*age.mfp.1 + age.beta.2*age.mfp.2 + size.beta*size.mfp +
    nodes.beta*nodes.mfp + grade.beta*grade.val + screen.beta*screen +
    her2.beta + ki67.beta + pr.beta

  # Non breast cancer mortality
  # Generate cumulative baseline other mortality

  time<-c(5,10,15)

  base.m.cum.oth <- exp(-6.052919 + (1.079863*log(time)) + (.3255321*time^.5))


  # Generate cumulative survival non-breast mortality
  #YW: cumulative non-breast cancer survival
  s.cum.oth <- exp(-exp(mi)*base.m.cum.oth)

  # Generate annual survival from cumulative survival
  #YW: cumulative non-breast cancer mortality
  m.cum.oth <- 1 - s.cum.oth

  # Breast cancer specific mortality
  # Generate cumulative baseline breast mortality
  if (er==1) {
    base.m.cum.br <- exp(0.7424402 - 7.527762/time^.5 - 1.812513*log(time)/time^.5)
  } else { base.m.cum.br <- exp(-1.156036 + 0.4707332/time^2 - 3.51355/time)
  }
  # Calculate the cumulative breast cancer survival
  s.cum.br <- exp(-exp(pi)*base.m.cum.br)
  m.cum.br <- 1 - s.cum.br

  # All cause mortality
  m.cum.all <- 1 - s.cum.oth*s.cum.br
  s.cum.all <- 100-100*m.cum.all

  # Proportion of all cause mortality that is breast cancer
  prop.br <- m.cum.br/(m.cum.br+m.cum.oth)
  prop.oth <- m.cum.oth/(m.cum.br+m.cum.oth)
  # Predicted cumulative breast specific mortality
  pred.m.br    <- prop.br*m.cum.all
  # Predicted cumulative non-breast cancer mortality
  pred.m.oth <- prop.oth*m.cum.all
  # Predicted cumulative all-cause mortality
  pred.all <- pred.m.br + pred.m.oth #pred.all=m.cum.all

  # Predicted all cause mortality
  comp_df$v23_all5[n] <- pred.all[1]
  comp_df$v23_all10[n] <- pred.all[2]
  comp_df$v23_all15[n] <- pred.all[3]


  #prognostic index, this is not related to time
  comp_df$v23_pi[n] <- pi #larger pi is associated with higher risks of events

}
```


# Calibration and discrimination
## PREDICT V2.2
```{r validity v22}
res_v22_all <- validity_Fun(data = comp_df, time = "timeyr", status = "delta", 
                         predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))

res_v22_ERpos <- validity_Fun(data = filter(comp_df, comp_df$er == 1), time = "timeyr", status = "delta", 
                         predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))

res_v22_ERneg <- validity_Fun(data = filter(comp_df, comp_df$er == 0), time = "timeyr", status = "delta", 
                         predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))

res_v22_HRposHER2neg <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRposHER2neg"), time = "timeyr", 
                                  status = "delta", predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))
res_v22_HRposHER2pos <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRposHER2pos"), time = "timeyr", 
                                  status = "delta", predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))
res_v22_HRnegHER2pos <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2pos"), time = "timeyr", 
                                  status = "delta", predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))
res_v22_HRnegHER2neg <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2neg"), time = "timeyr", 
                                  status = "delta", predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))


res_v22_HRnegHER2neg_gBRCA12m <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2neg" & comp_df$gBRCA12 == 1), time = "timeyr", 
                                  status = "delta", predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))

res_v22_HRnegHER2neg_gBRCA12wt <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2neg" & comp_df$gBRCA12 == 0), time = "timeyr", 
                                  status = "delta", predmort = c("v22_all5", "v22_all10", "v22_all15"), tmax = c(5, 10, 15))



res_v22 <- matrix(c(res_v22_all, res_v22_ERpos, res_v22_ERneg, res_v22_HRposHER2neg, res_v22_HRposHER2pos,
                    res_v22_HRnegHER2pos, res_v22_HRnegHER2neg, res_v22_HRnegHER2neg_gBRCA12m, res_v22_HRnegHER2neg_gBRCA12wt),
                      nrow = 9, byrow = T,
                      dimnames = list(c("All women", "ER-positive", "ER-negative", "HRposHER2neg", "HRposHER2pos", 
                                        "HRnegHER2pos", "HRnegHER2neg","gBRCA12m", "gBRCA12wt"), 
                                      c("No. women", "event_5","OE_5", "AUC_5", "event_10","OE_10", "AUC_10",
                                        "event_15","OE_15", "AUC_15"))) %>%
  as.data.frame(.)

res_v22
```

## PREDICT V2.3
```{r validity v23}
res_v23_all <- validity_Fun(data = comp_df, time = "timeyr", status = "delta", 
                         predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))

res_v23_ERpos <- validity_Fun(data = filter(comp_df, comp_df$er == 1), time = "timeyr", status = "delta", 
                         predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))

res_v23_ERneg <- validity_Fun(data = filter(comp_df, comp_df$er == 0), time = "timeyr", status = "delta", 
                         predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))

res_v23_HRposHER2neg <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRposHER2neg"), time = "timeyr", 
                                  status = "delta", predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))
res_v23_HRposHER2pos <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRposHER2pos"), time = "timeyr", 
                                  status = "delta", predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))
res_v23_HRnegHER2pos <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2pos"), time = "timeyr", 
                                  status = "delta", predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))
res_v23_HRnegHER2neg <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2neg"), time = "timeyr", 
                                  status = "delta", predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))


res_v23_HRnegHER2neg_gBRCA12m <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2neg" & comp_df$gBRCA12 == 1), time = "timeyr", 
                                  status = "delta", predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))

res_v23_HRnegHER2neg_gBRCA12wt <- validity_Fun(data = filter(comp_df, comp_df$intrsub == "HRnegHER2neg" & comp_df$gBRCA12 == 0), time = "timeyr", 
                                  status = "delta", predmort = c("v23_all5", "v23_all10", "v23_all15"), tmax = c(5, 10, 15))



res_v23 <- matrix(c(res_v23_all, res_v23_ERpos, res_v23_ERneg, res_v23_HRposHER2neg, res_v23_HRposHER2pos,
                    res_v23_HRnegHER2pos, res_v23_HRnegHER2neg, res_v23_HRnegHER2neg_gBRCA12m, res_v23_HRnegHER2neg_gBRCA12wt),
                      nrow = 9, byrow = T,
                      dimnames = list(c("All women", "ER-positive", "ER-negative", "HRposHER2neg", "HRposHER2pos", 
                                        "HRnegHER2pos", "HRnegHER2neg","gBRCA12m", "gBRCA12wt"), 
                                      c("No. women", "event_5","OE_5", "AUC_5", "event_10","OE_10", "AUC_10",
                                        "event_15","OE_15", "AUC_15"))) %>%
  as.data.frame(.)

res_v23
```




# Calibration plot
```{r cali plot}
calplot10_df_v22_all <- caliplot_df(data = comp_df, time= "timeyr", status = "delta", predmort = "v22_all10", tmax = 10)
calplot10_df_v22_ERpos <- caliplot_df(data = filter(comp_df, comp_df$er == 1), 
                                   time= "timeyr", status = "delta", predmort = "v22_all10", tmax = 10)
calplot10_df_v22_ERneg <- caliplot_df(data = filter(comp_df, comp_df$er == 0), 
                                   time= "timeyr", status = "delta", predmort = "v22_all10", tmax = 10)
```
## All women
```{r calibration all women, fig.width=12, fig.height=12}
caliplot_v22_all <-
  ggplot(calplot10_df_v22_all, aes(x = exp.x, y = obs.y)) +
  geom_point(color = "black", size = 3, shape = 15) +
  geom_line(color = "black", size = 1) +
  #abline! I used geomline instead of abline because I could specify the length of the abline
  geom_line(data = data.frame(x = c(0, 0.42), y = c(0, 0.42)), 
            aes(x, y), color = "black", linetype = "dashed") +
  geom_errorbar(aes(ymin = obs.lower, ymax = obs.upper),
                width = 0.02, color = "black", size = 0.7) +
  #geom_vline(xintercept = 0.12, color = "red", linetype = "dashed") + 
  xlim(0, 0.42) + ylim(0, 0.42) +
  xlab("Predicted 10-year all-cause mortality") +
  ylab("Observed 10-year all-cause mortality") +
  ggtitle("A. Calibration plot in all women") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 20), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 16),
        axis.line = element_line(color = "black", linewidth = 0.8),
        axis.text = element_text(size = 12)) 
caliplot_v22_all
```
## Women with ER-positive tumors
```{r calibration ERpos, fig.width=12, fig.height=12}
caliplot_v22_ERpos <-
  ggplot(calplot10_df_v22_ERpos, aes(x = exp.x, y = obs.y)) +
  geom_point(color = "black", size = 3, shape = 15) +
  geom_line(color = "black", size = 1) +
  #abline! I used geomline instead of abline because I could specify the length of the abline
  geom_line(data = data.frame(x = c(0, 0.42), y = c(0, 0.42)), 
            aes(x, y), color = "black", linetype = "dashed") +
  geom_errorbar(aes(ymin = obs.lower, ymax = obs.upper),
                width = 0.02, color = "black", size = 0.7) +
  #geom_vline(xintercept = 0.12, color = "red", linetype = "dashed") + 
  xlim(0, 0.42) + ylim(0, 0.42) +
  xlab("Predicted 10-year all-cause mortality") +
  ylab("Observed 10-year all-cause mortality") +
  ggtitle("B. Calibration plot in women with ER-positive tumors") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 20), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 16),
        axis.line = element_line(color = "black", linewidth = 0.8),
        axis.text = element_text(size = 12)) 
caliplot_v22_ERpos
```

## Women with ER-negative tumors
```{r calibration ERneg, fig.width=12, fig.height=12}
caliplot_v22_ERneg <-
  ggplot(calplot10_df_v22_ERneg, aes(x = exp.x, y = obs.y)) +
  geom_point(color = "black", size = 3, shape = 15) +
  geom_line(color = "black", size = 1) +
  #abline! I used geomline instead of abline because I could specify the length of the abline
  geom_line(data = data.frame(x = c(0, 0.42), y = c(0, 0.42)), 
            aes(x, y), color = "black", linetype = "dashed") +
  geom_errorbar(aes(ymin = obs.lower, ymax = obs.upper),
                width = 0.02, color = "black", size = 0.7) +
  #geom_vline(xintercept = 0.12, color = "red", linetype = "dashed") + 
  xlim(0, 0.42) + ylim(0, 0.42) +
  xlab("Predicted 10-year all-cause mortality") +
  ylab("Observed 10-year all-cause mortality") +
  ggtitle("C. Calibration plot in womwn with ER-negative tumors") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1.5, size = 20), 
        text = element_text(family = "Arial"),
        panel.background = element_rect(fill = "white"),
        axis.title = element_text(size = 16),
        axis.line = element_line(color = "black", linewidth = 0.8),
        axis.text = element_text(size = 12)) 
caliplot_v22_ERneg
```


# Session information
```{r sinfo}
sessionInfo()
```

